Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-08 doc/report/msc-report.tex
hard-wrapped a bunch of long lines. no content change.
improved abstract and introduction.

\noindent 
This thesis describes a data store for multi-dimensional time-series-based data. The data is aggregated over many units of measurements and across multiple dimensions. One of these dimensions is always time.

Zohmg solves the problem of summing, storing and serving measurements. These measurements may be taken from log files or similar sources. Zohmg stores only the summarized data, not the atomic measurements. Zohmg is optimized for speed of data retrieval.

@@

% TODO: den har texten passar battre i ett eget stycke/sektion, background reading style.

\chapterquote{Data warehouse is a repository of an organization's
electronically stored data. Data warehouses are designed to facilitate
reporting and analysis.}{Wikipedia}

The main actions of a data warehouse is extract, transform and load data,
although not necessarily in that order.

The classic thinking in data warehousing is the divide between facts and dimensons. The facts are measures categorized by dimensions. In OLAP, a sort of data warehouse, the OLAP cube is typically modelled in fact tables and dimension tables in a relational database. The novelty of our approach is to model OLAP cube-like aggregation data store on top of a bigtable-like data store.

As large scale web services grow they generate vast volumes of log data,
for instance Apache web logs. The ability to accurately analyse gathered
data is important to the evaluation, planning, and success of the web
service. Storing and analysing massive volumes of log data requires
special measures to be taken. From an end-user perspective the requirement
is an easy, complete and fast data analysing procedure. The technical
aspects of this requirement boils down to the following key points:
Visualising, storing and importing the data. The result is an
multidimensional information system which is subject to
OLAP\footnote{On-Line Analytical Processing}. \cite{olap_solutions}

All three requirements are depend on one another, what data and how it is
stored affects what can be visualised. Also, what data is imported---and
how it is analysed---affects how it can be stored and hence visualised.

@@

@@

% TODO> why is this Tool & BigTable-subsection called Dimensions? Disambiguation needed. -fredrik
\subsection{Dimensions}

@@

% TODO: not sure this is precisely a story about the implementation. putting it here for now.

@@

It'd be interesting to think a bit about how the data store could be fully elastic, i.e. not require any data set definition from the user and instead create column-families on the fly based on the dimensions output from the user's map.

\subsection{Multiple data sets}

>>>>

\noindent

This thesis describes a data store for multi-dimensional time-series-based
data. The data is aggregated over many units of measurements and across
multiple dimensions. One of these dimensions is always time.

In this thesis we present a data store called Zohmg. Zohmg is built on top of
HBase and uses Hadoop to run import jobs in parallel. It solves the problem of
summing, storing and serving measurements. These measurements may be taken
from log files or similar sources. Zohmg stores only the summarized data, not
the atomic measurements. The data store is optimized for speed of data
retrieval: one of the design goals was to serve data at mouse-click rate and
promote real-time data exploration.

Similar data stores exist: they generally use a relational database system as
their underlying data store. The novelty of our approach is to model OLAP
cubes on top of a bigtable-like data store.

@@

Any organization will eventually wish to evaluate and analyze various
characteristics of the service and operations they manage. These services can
be anything from web sites to theme parks, and the evaluation might include
characteristics such as the number of visitors per day, and the average time a
theme park guest spends queueing for a particular ride.

Inevitably, these services and operations will generate a log for later
analysis. As the services grow in size, so will the data sets they generate grow.

An important property of these data sets is that they generally have a regular
structure and are represented in independent records. The records can be
broken up without losing any of their meaning and are as such very well suited
for parallel analysis. The data sets are not fit to store using traditional
database techniques, if only for the fact that they are too large to fit in a
relational database.

@@

\section{Other data warehouses}

A data warehouse is a repository of an organization's
electronically stored data. Data warehouses are designed to facilitate
reporting and analysis. \cite{wikipedia}

The core actions of a data warehouse is to load, transform and extract data.
The classic thinking in modelling data warehouses is to divide data into facts
and dimensons. The facts are measures categorized by dimensions.

In OLAP\footnote{On-Line Analytical Processing}, which is a kind of data
warehouse, the data is thought of as being structured as a cube. The cube is
typically modelled in a realational database with fact tables and dimension
tables, upon which the system performs joins to extract data.
\cite{olap_solutions}

@@

% TODO> why is this Tool & BigTable-subsection called Dimensions?
% Disambiguation needed. -fredrik
\subsection{Dimensions}

@@

% TODO: not sure this is precisely a story about the implementation.
% putting it here for now.

@@

It'd be interesting to think a bit about how the data store could be fully
elastic, i.e. not require any data set definition from the user and instead
create column-families on the fly based on the dimensions output from the
user's map.
\subsection{Multiple data sets}
